---
title: "Homework 3 Air Quality Risk Assessment"
author: "Rebeca Adam"
format: html
editor: visual
---

```{r}
library(tidyverse)
```

## Part 1: Building the Dataset

```{r}
day = 1:365

pm25 = rnorm(n = 365, mean = 14, sd = 8)

pm25 = ifelse(pm25 < 0, 0, pm25)

pm25

```


```{r}
season = rep(NA, times = 365)

for (i in 1:365) {
  if (i <= 90) {
    season[i] = "winter"
  } else if (i <= 180) {
    season[i] = "spring"
  } else if (i <= 270) {
    season[i] = "summer"
  } else {
    season[i] = "fall"
  }
}

season = as.factor(season)

season
```


```{r}
airquality = data.frame(day = day, pm25 = pm25, season = season)

airquality$risk = NA

airquality
```

#Justification

I used a normal distribution (rnorm) to simulate PM2.5 concentration of 14 ug/m^3 and a standard deviation of 8 to simulate moderate urban air quality conditions with a realistic day-to-day variablility. The EPA defines "good" air quality around >= 12ug/m^3, and this analysis seeks to assess a moderate scenario while producing enough variation to evaluate.








