---
title: "Homework 3 Air Quality Risk Assessment"
author: "Rebeca Adam"
format: html
editor: visual
---

```{r}
library(tidyverse)
```

## Part 1: Building the Dataset

```{r}
day = 1:365

pm25 = rnorm(n = 365, mean = 19, sd = 10)

pm25 = ifelse(pm25 < 0, 0, pm25)

pm25

```


```{r}
season = rep(NA, times = 365)

for (i in 1:365) {
  if (i <= 90) {
    season[i] = "winter"
  } else if (i <= 180) {
    season[i] = "spring"
  } else if (i <= 270) {
    season[i] = "summer"
  } else {
    season[i] = "fall"
  }
}

season = as.factor(season)

season
```


```{r}
airquality = data.frame(day = day, pm25 = pm25, season = season)

airquality$risk = NA

airquality
```

#Justification

I used a normal distribution (rnorm) to simulate PM2.5 concentration of 19 ug/m^3 and a standard deviation of 10 to simulate moderate urban air quality conditions with a realistic day-to-day variablility. The Indoor Air Hygiene Institute defines "good" air quality around >= 12ug/m^3, and this analysis seeks to assess a moderate scenario while producing enough variation to evaluate. If levels go above 35ug/m^3 during a 24-hour period, exposure is considered unhealthy


## Part 2 - Risk Loop

```{r}
# Initializing
current_streak = 0
max_unhealthy_streak = 0
```


```{r}
for (i in 1:nrow(airquality)) {
  if (airquality$pm25[i] < 12) {
    airquality$risk[i] = "good"
    current_streak = 0
  } else if (airquality$pm25[i] <= 35) {
    airquality$risk[i] = "moderate"
    current_streak = 0
  } else {
    airquality$risk[i] = "unhealthy"
    current_streak = current_streak + 1
    if (current_streak > max_unhealthy_streak) {
      max_unhealthy_streak = current_streak
    }
  }
}
```











